---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bypass

{bypass} provides a way to use the default behavior of internal generics, bypassing
the S3 methods. This makes working with objects at the low level much easier
and safer.

We provide counterparts to existing functions:

| {base} | {bypass} |
|------|------|
|  c | .c |
|  dim | .dim |
|  dim<- | .dim<- |
|  $ | .dollar |
|  $<- | .dollar<- |
|  lapply | .lapply |
|  length | .length |
|  lengths | .lengths |
|  names | .names |
|  names<- | .names<- |
|  sapply | .sapply |
|  [<- | .subset<- |
|  [[<- | .subset2<- |
|  unlist | .unlist |

Note that base R already has `.subset()` and `.subset2()` as low level counterparts
to `[` and `[[`. check also {rlang} for low level versions of `is.symbol()`,
`is.na()` etc

Additionally the functions `with_bypass()` and `local_bypass()` allow you
to locally type the base versions and get the bypass behavior. This is especially
useful for `[<-` and `[[<-`.

## Installation

You can install the development version of bypass like so:

``` r
pak::pak("cynkra/bypass")
```

## Example 1

```{r example}
library(bypass)
## from ?bibentry
rref <- bibentry(
   bibtype = "Manual",
   title = "R: A Language and Environment for Statistical Computing",
   author = person("R Core Team"),
   organization = "R Foundation for Statistical Computing",
   address = "Vienna, Austria",
   year = 2014,
   url = "https://www.R-project.org/"
   )

# The S3 method makes the object hard to manipulate
identical(rref[[1]][[1]], rref)

# Base R helps here
.subset2(rref, c(1, 1))

# though it looks bad when we want to combine indices types
.subset2(rref, 1) |> .subset2("title")

# with {bypass} we can use the standard syntax
with_bypass(rref[[1]][["title"]])

# `local_bypass()` is meant to be used in functions and is probably what
# you'll want to use
fun <- function(x) {
  local_bypass()
  x[[1]]$title <- "!!!!!!!!!!!!!!!!"
  x
}
fun(rref)
```

## Example 2

```{r}
x <- structure(list(a = 1, b = 2), class = "foo")
c.foo <- function(...) 42
dim.foo <- function(...) 42
`$.foo` <- function(...) 42
lapply.foo <- function(...) 42
length.foo <- function(...) 42
lengths.foo <- function(...) 42
names.foo <- function(...) 42
sapply.foo <- function(...) 42
`[.foo` <- function(...) 42
`[[.foo` <- function(...) 42
unlist.foo <- function(...) 42
  
c(x)
.c(x)

dim(x)
.dim(x)

x$a
.dollar(x, a)

lapply(x, identity)
.lapply(x, identity)

sapply(x, identity)
.sapply(x, identity)

length(x)
.length(x)

names(x)
.names(x)

unlist(x)
.unlist(x)

# NOTE: `local_bypass()` should normally be called in a function so we don't 
# change the global state, here for demo purposes:
local_bypass()
c(x)
dim(x)
x$a
lapply(x, identity)
sapply(x, identity)
length(x)
names(x)
unlist(x)
```

