---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# bypass

{bypass} provides a way to use the default behavior of internal generics, bypassing
the S3 methods. This makes working with objects at the low level much easier
and safer.

We provide counterparts to existing functions:

| {base} | {bypass} |
|------|------|
|  c | .c |
|  dim | .dim |
|  dim<- | .dim<- |
|  $ | .dollar |
|  $<- | .dollar<- |
|  lapply | .lapply |
|  length | .length |
|  lengths | .lengths |
|  names | .names |
|  names<- | .names<- |
|  sapply | .sapply |
|  [< | .subset1 |
|  [<- | .subset1<- |
|  [[<- | .subset2<- |
|  unlist | .unlist |

Note that base R already has `.subset2()` as a low level counterpart
to `[[`. `.subset1()` is different from `.subset()` because it doesn't lose attributes
other than names. `.dollar()` is a wrapper around `.subset2()` so we also lose
the partial matching of the original `$` function. 
 
3 additional functions are provided :

*  `with_bypass()` allows you to type the base versions in the `expr`
  argument and get the bypass behavior
* `local_bypass()` does the same but locally for the function in which it's called
* `global_bypass()` does this for a full package, it's meant to be used in `.onLoad()`

These are especially useful useful for `[<-` and `[[<-` which are tricky to use
at the low level, as they need a lot of unclassing/reclassing.

## Installation

You can install the development version of bypass like so:

``` r
pak::pak("cynkra/bypass")
```

## Example 1

At the high level POSIXlt objects look like simple vectors, and behave that way

```{r example}
library(bypass)
x <- as.POSIXlt(c("2024-01-01", "2024-01-02"))
length(x)
names(x)
```

But in fact "POSIXlt" objects are not atomic vectors but named lists of vectors,
S3 methods, for `length()`, `names()`, and more, are defined so we can treat them
just like vectors. 

Here's another way we might have used to define the above object:

```{r}
list(
  sec = c(0, 0),
  min = c(0L, 0L),
  hour = c(0L, 0L),
  mday = 1:2,
  mon = c(0L, 0L),
  year = c(124L, 124L),
  wday = 1:2,
  yday = 0:1,
  isdst = c(0L, 0L),
  zone = c("CET", "CET"),
  gmtoff = c(NA_integer_, NA_integer_)
) |>
  structure(class = c("POSIXlt", "POSIXt"), tzone = c("", "CET", "CEST"), balanced = TRUE)
```

The `.length()` and `.names()` functions can be used to access the low level
length and names, these are roughly equivalent to `length(unclass(x))` and 
`names(unclass(x))` respectively, with special cases for environments.

```{r}
.length(x)
.names(x)
```

The functions `with_bypass()`, `local_bypass()` and `global_bypass()` provide
different ways to use the native syntax rather than dotted counterparts.

```{r}
# with_bypass
with_bypass(names(x))

# local_bypass
fun <- function() {
  local_bypass()
  names(x)
}
fun()

# global_bypass (in a package)
.onLoad <- function(libname, pkgname) {
  bypass::global_bypass(asNamespace(pkgname))
}
# then names() will behave like .names() in the whole package
```

Low level replacement is one of the most useful features.

```{r}
# without bypass
x <- as.POSIXlt(c("2024-01-01", "2024-01-02"))
cl <- class(x)
x <- unclass(x)
x$year <- c(120L, 121L)
class(x) <- cl
x

# with bypass
x <- as.POSIXlt(c("2024-01-01", "2024-01-02"))
with_bypass({
  x$year <- c(120L, 121L)
})
x
```

In case of replacement of nested elements the difference between both approaches
will be even bigger.

## Example 2

```{r}
x <- structure(list(a = 1, b = 2), class = "foo")
c.foo <- function(...) 42
dim.foo <- function(...) 42
`$.foo` <- function(...) 42
length.foo <- function(...) 42
lengths.foo <- function(...) 42
names.foo <- function(...) 42
sapply.foo <- function(...) 42
`[.foo` <- function(...) 42
`[[.foo` <- function(...) 42
unlist.foo <- function(...) 42
  
c(x)
.c(x)

dim(x)
.dim(x)

x$a
.dollar(x, a)

lapply(x, identity)
.lapply(x, identity)

sapply(x, identity)
.sapply(x, identity)

length(x)
.length(x)

names(x)
.names(x)

unlist(x)
.unlist(x)

# NOTE: `local_bypass()` should normally be called in a function so we don't 
# change the global state, here for demo purposes:
local_bypass()
c(x)
dim(x)
x$a
lapply(x, identity)
sapply(x, identity)
length(x)
names(x)
unlist(x)
```

